Quantifying stochastic noise in cultured\\ circadian reporter cells

### 2015-03-11 main_text.tex

\begin{figure}[tbp]
  \begin{center}
    \includegraphics[width=0.8\textwidth]{figures/pdfs/noise_ts_and_boot.pdf}
  \end{center}
  \caption{{\bfseries Single-cell bioluminescence recordings show that higher stochastic noise results in faster damping at the population-level.}
  Data on the bioluminescence of single-cell fibroblasts was taken from Leise {\itshape et al.,} 2012 \cite{Leise2012}.
({\bfseries A}) Cells were sorted into two groups depending on their degree of stochastic noise. An example trace from each of the two groups is shown, demonstrating different levels of noise present in the dataset.
({\bfseries B}) After artificially synchronizing each cell, we calculate averaged bioluminescence rhythms of each group (solid lines). A damped sinusoid fit to both groups reveals a difference in damping rate, demonstrated by fitted envelope functions ($\pm A\exp{-dt}$, dashed lines).
({\bfseries C}) The observed absolute difference in damping rate was shown to be significant ($p = 0.0264$) using $10,000$ bootstrap trials.}
\label{fig:fibroblast_noise}
\end{figure}

@@

\begin{figure}[tbp]
  \begin{center}
    \includegraphics[width=0.8\textwidth]{figures/pdfs/fitted_parameters.pdf}
  \end{center}
  \caption{{\bfseries Distributions in fitted parameters for the genome-wide siRNA screen.} Distributions in robust z-scores closely resemble normal distributions. For all parameters, the region of highest density is consistent between the control and perturbed populations, indicating many perturbations do not appreciably change clock dynamics. } 
\label{fig:fit_distributions}
\end{figure}

@@

\begin{figure}[tbp]
  \begin{center}
    \includegraphics[width=0.65\textwidth]{figures/pdfs/outliers.pdf}
  \end{center}
  \caption{{\bfseries Effects of siRNA knockdowns on amplitude and damping rate.}
  Clock robustness is a function of both amplitude and damping rate.
  Distributions in amplitude and damping rate for control wells ({\bfseries A}) or perturbed wells ({\bfseries B}) indicate that perturbations tend to shift the clock towards regions of higher damping rate or lower amplitude. Green dots in each figure indicates the mean of the control population.
({\bfseries C}) Averaged effect of siRNA knockdown after grouping the perturbed population by Gene ID. Only those genes which were significantly different from the control distribution are shown (Hotelling's $T^2$ test, $\alpha = 0.01$).
({\bfseries D}) Radial histogram of the significant gene perturbations shown in (C). The area of each slice is proportional to the frequency of perturbations away from the mean in that direction. Very few gene knockdowns result in both higher amplitude and lower damping rate (red slices, lower right quadrant).
}
\label{fig:outlier_dist}
\end{figure}

@@

\begin{figure}[tbp]
  \begin{center}
    \includegraphics[width=0.5\textwidth]{figures/pdfs/small_molecule_r2.pdf}
  \end{center}
  \caption{{\bfseries Fit quality for the dose-dependent small molecule screens.} The bioluminescence rhythms in both reporter systems were well-described by a damped sinusoid. As the molecules were toxic to the reporter cells at high concentrations, fit quality declined with increasing dosage. Only fits with $R^2 > 0.8$ were kept for further analysis.}
\label{fig:small_molecule_r2}
\end{figure}

\begin{figure}[tbp]
  \begin{center}
    \includegraphics[width=0.5\textwidth]{figures/pdfs/volume_calibration.pdf}
  \end{center}
  \caption{{\bfseries Calibration curve for fitting the volume parameter to experimental data.} The model's volume parameter was linearly related to the population-averaged damping rate on a log-log scale. Error bars represent the standard error of the mean, calculated by $10$ independent replicates for each volume. Points shown in gray had an average $R^2 < 0.9$ and were excluded from the linear regression. Solid and dashed grey lines indicate the mean and $95\%$ confidence intervals of the linear regression, respectively.}
\label{fig:vol_calibration}
\end{figure}

\begin{figure}[tbp]
  \begin{center}
    \includegraphics[width=0.8\textwidth]{figures/pdfs/zhang_plates.pdf}
  \end{center}
  \caption{{\bfseries Plate-to-plate variation of fitted parameters in the Zhang {\itshape et al.,} 2009 genome-wide siRNA screen.} Dots indicate the median of each plate, with lines extending from the $5^\text{th}$ to $95^\text{th}$ percentile. While parameter fits were of similar magnitude for all plates, some inconsistencies were present. In order to accurately compare perturbations and controls between plate experiments, we normalized fitted parameters on a plate-by-plate basis.}
\label{fig:plate_variation}
\end{figure}

\begin{figure}[tbp]
  \begin{center}
    \includegraphics[width=0.8\textwidth]{figures/pdfs/zhang_wells.pdf}
  \end{center}
  \caption{{\bfseries Well-to-well variation in the Zhang {\itshape et al.,} 2009 genome-wide siRNA screen.} Similar to Figure~\ref{fig:plate_variation}, dots indicate the median value of each parameter in each well, with lines extending from the $5^\text{th}$ to $95^\text{th}$ percentile. Well position did not seem to affect the fitted values, particularly in the middle regions which contained the siRNA knockdown library. Wells on either end showed significant variation, but these are likely due to the spotting of long- and short-period controls in the same position on each plate.}
\label{fig:well_variation}
\end{figure}

>>>>

\begin{figure}[tbp]
  \begin{center}
    % \includegraphics[]{figures/pdfs/Fig1.pdf}
  \end{center}
  \caption{{\bfseries Single-cell bioluminescence recordings show that higher stochastic noise results in faster damping at the population-level.}
  Data on the bioluminescence of single-cell fibroblasts was taken from Leise {\itshape et al.,} 2012 \cite{Leise2012}.
({\bfseries A}) Cells were sorted into two groups depending on their degree of stochastic noise. An example trace from each of the two groups is shown, demonstrating different levels of noise present in the dataset.
({\bfseries B}) After artificially synchronizing each cell, we calculate averaged bioluminescence rhythms of each group (solid lines). A damped sinusoid fit to both groups reveals a difference in damping rate, demonstrated by fitted envelope functions ($\pm A\exp{-dt}$, dashed lines).
({\bfseries C}) The observed absolute difference in damping rate was shown to be significant ($p = 0.0264$) using $10,000$ bootstrap trials.}
\label{fig:fibroblast_noise}
\end{figure}

@@

\begin{figure}[tbp]
  \begin{center}
    % \includegraphics[]{figures/pdfs/Fig5.pdf}
  \end{center}
  \caption{{\bfseries Distributions in fitted parameters for the genome-wide siRNA screen.} Distributions in robust z-scores closely resemble normal distributions. For all parameters, the region of highest density is consistent between the control and perturbed populations, indicating many perturbations do not appreciably change clock dynamics. } 
\label{fig:fit_distributions}
\end{figure}

@@

\begin{figure}[tbp]
  \begin{center}
    % \includegraphics[]{figures/pdfs/Fig6.pdf}
  \end{center}
  \caption{{\bfseries Effects of siRNA knockdowns on amplitude and damping rate.}
  Clock robustness is a function of both amplitude and damping rate.
  Distributions in amplitude and damping rate for control wells ({\bfseries A}) or perturbed wells ({\bfseries B}) indicate that perturbations tend to shift the clock towards regions of higher damping rate or lower amplitude. Green dots in each figure indicates the mean of the control population.
({\bfseries C}) Averaged effect of siRNA knockdown after grouping the perturbed population by Gene ID. Only those genes which were significantly different from the control distribution are shown (Hotelling's $T^2$ test, $\alpha = 0.01$).
({\bfseries D}) Radial histogram of the significant gene perturbations shown in (C). The area of each slice is proportional to the frequency of perturbations away from the mean in that direction. Very few gene knockdowns result in both higher amplitude and lower damping rate (red slices, lower right quadrant).
}
\label{fig:outlier_dist}
\end{figure}

@@

\begin{figure}[tbp]
  \begin{center}
    % \includegraphics[]{figures/pdfs/FigS1.pdf}
  \end{center}
  \caption{{\bfseries Fit quality for the dose-dependent small molecule screens.} The bioluminescence rhythms in both reporter systems were well-described by a damped sinusoid. As the molecules were toxic to the reporter cells at high concentrations, fit quality declined with increasing dosage. Only fits with $R^2 > 0.8$ were kept for further analysis.}
\label{fig:small_molecule_r2}
\end{figure}

\begin{figure}[tbp]
  \begin{center}
    % \includegraphics[]{figures/pdfs/FigS2.pdf}
  \end{center}
  \caption{{\bfseries Calibration curve for fitting the volume parameter to experimental data.} The model's volume parameter was linearly related to the population-averaged damping rate on a log-log scale. Error bars represent the standard error of the mean, calculated by $10$ independent replicates for each volume. Points shown in gray had an average $R^2 < 0.9$ and were excluded from the linear regression. Solid and dashed grey lines indicate the mean and $95\%$ confidence intervals of the linear regression, respectively.}
\label{fig:vol_calibration}
\end{figure}

\begin{figure}[tbp]
  \begin{center}
    % \includegraphics[]{figures/pdfs/FigS3.pdf}
  \end{center}
  \caption{{\bfseries Plate-to-plate variation of fitted parameters in the Zhang {\itshape et al.,} 2009 genome-wide siRNA screen.} Dots indicate the median of each plate, with lines extending from the $5^\text{th}$ to $95^\text{th}$ percentile. While parameter fits were of similar magnitude for all plates, some inconsistencies were present. In order to accurately compare perturbations and controls between plate experiments, we normalized fitted parameters on a plate-by-plate basis.}
\label{fig:plate_variation}
\end{figure}

\begin{figure}[tbp]
  \begin{center}
    % \includegraphics[]{figures/pdfs/FigS4.pdf}
  \end{center}
  \caption{{\bfseries Well-to-well variation in the Zhang {\itshape et al.,} 2009 genome-wide siRNA screen.} Similar to Figure~\ref{fig:plate_variation}, dots indicate the median value of each parameter in each well, with lines extending from the $5^\text{th}$ to $95^\text{th}$ percentile. Well position did not seem to affect the fitted values, particularly in the middle regions which contained the siRNA knockdown library. Wells on either end showed significant variation, but these are likely due to the spotting of long- and short-period controls in the same position on each plate.}
\label{fig:well_variation}
\end{figure}
